### 2023-08-19 docs/presentation/presentation.tex
Update presentation

\documentclass{beamer}
\usepackage{amsfonts,amsmath,oldgerm}
\usepackage{bytefield}
\usepackage{listings}

% pympress notes page config
\iffalse
	\setbeamertemplate{note page}
	{
		\insertslideintonotes{0.01}
		\rule{\textwidth}{0.1pt}
		%\color{blue} \scriptsize
		\insertnote
	}
	\setbeameroption{show notes on second screen}
\fi

% adjust gigantic tabs in code blocks, not actually rust format
\lstdefinestyle{rust}{
	tabsize=2
}
\lstset{style=rust}

\usetheme{sintef}
\usefonttheme[onlymath]{serif}
\titlebackground*{assets/background}

% needed this a bunch so made a macro for myself
\newcommand{\rot}[1]{\rotatebox{90}{#1}}

\title{Implementing XBF: An Efficient Self-Describing Binary Format}
\author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}

\begin{document}
\maketitle

\section{Introduction}

\begin{frame}{Web Data Interchange Formats}
	\begin{itemize}
		\item Human-Readable and Self-Describing
		      \note[item]{Most formats on the web are human-readable and self-describing}
		      \note[item]{Self describing means they include metadata to describe themselves, and don't require an external schema}
		      \note[item]{Human readable means plain text, ASCII or utf-8}
		\item Most common: JSON and XML
		      \note[item]{The most common of these formats is JSON and XML}
		\item Alternatives: CSV, MessagePack, CBOR
		      \note[item]{There are other human-readable formats such as CSV}
		      \note[item]{And others, such as CBOR (concise binary object representation), and MessagePack that are still self-describing, but use a binary format instead}
	\end{itemize}
\end{frame}

\begin{frame}{Shortcomings of Current Formats}
	\begin{itemize}
		\item Repeated sending of metadata.
		\item Inefficiency of plain text.
		      \note[item]{Expensive conversion between plain text and binary}
		      \note[item]{Requires more bytes to store the same data}
		\item Use of big-endian.
		      \note[item]{Most modern processors (x86 and ARM) are little-endian}
		      \note[item]{Both CBOR and MessagePack are big-endian}
	\end{itemize}
	% XML from paper here
\end{frame}

\begin{frame}[fragile]{Repeated Metadata}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\note[item]{XML example where a huge percent of the actual data sent is metadata}
			\begin{lstlisting}[language=XML]
<person>
	<name>John Jackson</name>
	<age>25</age>
</person>
\end{lstlisting}
		\end{column}
		\begin{column}{0.5\textwidth}
			\begin{table}
				\caption{XML Space Utilization}
				\center
				\begin{tabular}{|c|c|c|}
					\hline
					\textbf{} & \textbf{Bytes Used} & \textbf{Percent} \\
					\hline
					Metadata  & 45                  & 76\%             \\
					Data      & 14                  & 24\%             \\
					\hline
					Total     & 59                  & 100              \\
					\hline
				\end{tabular}
			\end{table}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Plain Text Example}
	\begin{table}
		\caption{Sending A 32-bit Unsigned Integer}
		\begin{center}
			\begin{tabular}{|c|c|c|}
				\hline
				\textbf{Number} & \textbf{ASCII Bytes} & \textbf{Binary Bytes} \\
				\hline
				1               & 1                    & 4                     \\
				9999            & 4                    & 4                     \\
				7654321         & 7                    & 4                     \\
				\hline
			\end{tabular}
		\end{center}
	\end{table}
\end{frame}

\begin{frame}[fragile]{Big Endian Example: 32-bit Unsigned Integer}
	\begin{center}
		\textbf{Number To Send: 1027} \\
	\end{center}
	\begin{columns}
		\begin{column}{0.3\textwidth}
			\begin{bytefield}{16}
				\wordbox{1}{Sender (Little Endian)} \\
				\bitboxes{4}{{3} {4} {0} {0}} \\
			\end{bytefield}
		\end{column}
		\begin{column}{0.3\textwidth}
			\begin{bytefield}{16}
				\wordbox{1}{On the Wire (Big Endian)} \\
				\bitboxes{4}{{0} {0} {4} {3}} \\
			\end{bytefield}
		\end{column}
		\begin{column}{0.3\textwidth}
			\begin{bytefield}{16}
				\wordbox{1}{Recipient (Little Endian)} \\
				\bitboxes{4}{{3} {4} {0} {0}} \\
			\end{bytefield}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Improvements For a New Format}
	\begin{itemize}
		\item When possible, send metadata only a single time (mainly for homogenous lists).
		\item Allow sending metadata to be optional, such as if a client already has it.
		      \note[item]{Keep metadata and data separate, not coupled together}
		\item Utilize little-endian to eliminate endian conversions on common architectures.
		\item Prioritize a simple type system.
		      \note[item]{Hopefully makes the implementation simpler}
	\end{itemize}
\end{frame}

\section{Format Design}

\begin{frame}{Primitives and Primitive Metadata}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\begin{itemize}
				\item Boolean
				      \note[item]{True of False, one byte}
				      \note[item]{All integers are little-endian}
				      \note[item]{Integer types are equivalent to the types in stdint.h C headers}
				\item U8, U16, U32, U64, U128, U256
				      \note[item]{Unsigned Integer Types}
				      \begin{itemize}
					      \item Equivalent to uint*\_t types in C
				      \end{itemize}
				\item I8, I16, I32, I64, I128, I256
				      \note[item]{Signed Integer Types}
				      \begin{itemize}
					      \item Equivalent to int*\_t types in C
				      \end{itemize}
				\item F32, F64
				      \note[item]{Floating Point Types}
				      \begin{itemize}
					      \item IEEE 754 floating point numbers
					      \item Equivalent to float and double in C
				      \end{itemize}
				\item Bytes  (Non-UTF-8 Bytes)
				      \note[item]{String but doesn't have to be UTF-8}
				\item String (UTF-8 Bytes)
				      \note[item]{Sends length as u64 followed by UTF-8 bytes}
			\end{itemize}
		\end{column}
		\begin{column}{0.25\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Value} \\
				\hline
				Boolean       & 0              \\
				U8            & 1              \\
				U16           & 2              \\
				U32           & 3              \\
				U64           & 4              \\
				U128          & 5              \\
				U256          & 6              \\
				\hline
			\end{tabular}
		\end{column}
		\begin{column}{0.25\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Value} \\
				\hline
				I8            & 7              \\
				I16           & 8              \\
				I32           & 9              \\
				I64           & 10             \\
				I128          & 11             \\
				I256          & 12             \\
				F32           & 13             \\
				F64           & 14             \\
				Bytes         & 15             \\
				String        & 16             \\
				\hline
			\end{tabular}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Primitive Example}
	\begin{itemize}
		\item 16-Bit Unsigned Integer: 1024 \\
		      \begin{bytefield}{24}
			      \bitbox{8}{Disc} & \bitbox{16}{1024} \\
			      \bitbox{8}{2} & \bitboxes{8}{04} \\
		      \end{bytefield}
		\item String: "hello" \\
		      \begin{bytefield}[bitheight=\widthof{"Disc"}]{28}
			      \bitbox{2}{\rot{Disc}} & \bitbox{16}{length} & \bitboxes{2}{hello} \\
			      \bitboxes{2}{{16} {2} {0} {0} {0} {0} {0} {0} {0} {\rot{104}} {\rot{101}} {\rot{108}} {\rot{108}} {\rot{111}}}
		      \end{bytefield}

	\end{itemize}
\end{frame}

\begin{frame}{Vectors (Homogenous Lists)}
	Values
	\begin{itemize}
		\item Homogenous list of values that has a known length.
		\item Length is sent as a part of the type, not in metadata.
		\item Inner type Information is \emph{not} sent as a part of the value.
	\end{itemize}

	Metadata
	\begin{itemize}
		\item Single byte discriminant value is sent (String + 1).
		\item Followed by internal type metadata (can continue recursively).
		\item Length is \emph{not} sent as a part of the metadata.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Vector Example}
	\begin{center}
		Vector of 16-Bit Unsigned Integers: [42, 1024] \vfill \break

		DV = Discriminant of a Vector

		D16 = Discriminant of a 16-Bit Unsigned Integer \vfill \break

		\begin{bytefield}{26}
			\bitbox{2}{\rot{DV}} & \bitbox{2}{\rot{D16}} & \bitbox{16}{length} & \bitboxes{4}{{42} {1024}} \\
			\bitboxes{2}{{17} {2} {2} {0} {0} {0} {0} {0} {0} {0} {42} {0} {0} {4}}
		\end{bytefield}
	\end{center}
\end{frame}

\begin{frame}{Structs (Aggregate Types)}
	\begin{itemize}
		\item	An aggregate type containing a name as well as named fields.
		\item May not contain duplicate field names.
		\item Fields are sent in sequence in the order they are listed in metadata.
		\item Serialized value does \emph{not} include:
		      \begin{itemize}
			      \item Struct name.
			      \item Number of fields.
			      \item Field names.
			      \item Field types.
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Struct Metadata}
	\begin{itemize}
		\item Single byte discriminant value is sent (Vector + 1).
		\item Followed by Struct name.
		\item Next, number of fields as an unsigned 16-bit integer.
		      \note[item]{A 16-bit integer is used because it's very unlikely for a struct of more than 65 thousand fields to be necessary.}
		\item Last, pairs of field names and types.
		      \begin{itemize}
			      \item May continue recursively with nested types of Structs or Vectors.
		      \end{itemize}
		\item All names are sent with the same format as primitive Strings.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Struct Metadata Example}
	\begin{columns}
		\begin{column}{0.65\textwidth}
			\center
			\begin{bytefield}{32}
				\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
				\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DF}} \\
				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DL}} \\
				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
			\end{bytefield}
		\end{column}
		\begin{column}{0.35\textwidth}
			\begin{lstlisting}
struct Point {
	first: i8,
	last: i8
}
\end{lstlisting}
			\hfill \break
			DS = Struct Discriminant

			DF = Field "first" discriminant

			DL = Field "last" discriminant
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Struct Value Example}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\center
			\begin{bytefield}{12}
				\bitbox{6}{first} & \bitbox{6}{last} \\
				\bitbox{6}{42} & \bitbox{6}{251} \\
			\end{bytefield}
		\end{column}
		\begin{column}{0.5\textwidth}
			\center
			\begin{lstlisting}
let my_point = Point {
	first: 42,
	last: -5,
};
			\end{lstlisting}
		\end{column}
	\end{columns}
\end{frame}

\section{Evaluation}

\begin{frame}{Encoder and Decoder Reference Implementation}
	\begin{itemize}
		\item Written in Rust.
		      \note[item]{systems level programming language similar to C and C++, strong memory safety guarantees}
		\item Extensively Tested.
		      \note[item]{100 percent code coverage}
		      \note[item]{not going to talk a ton about the reference implementation}
		\item Majority of time was spent on this implementation.
	\end{itemize}
\end{frame}

\begin{frame}{Performance Test}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			Server
			\begin{itemize}
				\item Downloads and parses 1 year of Sony stock history.
				\item 
			\end{itemize}
		\end{column}
		\begin{column}{0.5\textwidth}
				\item For each data format the client:
				      \begin{itemize}
					      \item Client sends 100 requests to the server.
					      \item Records average time to receive a response.
					      \item Records number of bytes read.
				      \end{itemize}
		\end{column}
	\end{columns}
\end{frame}

\section{Results}

\begin{frame}{Average Time and Bytes Read}
	\center
	\begin{tabular}{|c|c|c|c|c|}
		\hline
		\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\ 
		\hline                                                                                                                    
		CSV             & 18.93                  & 16,411              & 1,823                     & 11.1                      \\ 
		MessagePack     & 11.22                  & 15,565              & 977                       & 6.22                      \\ 
		CBOR            & 16.95                  & 25,507              & 10,919                    & 42.8                      \\ 
		JSON            & 21.91                  & 31,180              & 16,592                    & 53.2                      \\ 
		XML             & 21.87                  & 43,699              & 29,111                    & 66.6                      \\ 
		XBF             & 11.32                  & 14,686              & 98                        & 0.67                      \\ 
		\hline
	\end{tabular}
\end{frame}

\begin{frame}{Overhead}
	\center
	\begin{tabular}{|c|c|c|}
		\hline
		\textbf{Format} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\
		\hline
		CSV             & 1,823                     & 11.1                      \\
		MessagePack     & 977                       & 6.22                      \\
		CBOR            & 10,919                    & 42.8                      \\
		JSON            & 16,592                    & 53.2                      \\
		XML             & 29,111                    & 66.6                      \\
		XBF             & 98                        & 0.67                      \\
		\hline
	\end{tabular}
\end{frame}

\backmatter[notitle]
\end{document}